\documentclass{tsp-short}

\usepackage{amsthm}

\newtheorem{assumption}{Assumption}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}{Lemma}[section]
\newtheorem{corollary}{Corollary}[section]
\newtheorem{proposition}{Proposition}[section]

\theoremstyle{remark}
\newtheorem{remark}{Remark}[section]

\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]
\newtheorem{example}{Example}[section]

\newcommand{\R}{\mathbb{R}}
\newcommand{\cP}{\mathcal{P}}
\newcommand{\E}{\mathbb E}
\newcommand{\supp}{\mathrm{supp}}
\newcommand{\op}{\mathrm{op}}
\newcommand{\HS}{\mathrm{HS}}
\DeclareMathOperator{\Ent}{Ent}
\DeclareMathOperator{\Var}{Var}
\DeclareMathOperator{\Lip}{Lip}

\usepackage[unicode,hidelinks]{hyperref}
\usepackage[nameinlink,noabbrev]{cleveref} % for cref

\crefname{assumption}{Assumption}{Assumptions}
\Crefname{assumption}{Assumption}{Assumptions}

\title[Interacting SDEs]{Functional Inequalities and Synchronization for the measure-valued solutions of Stochastic Differential Equations with Interaction}

\author{Kyrylo Kuchynskyi}
\address{Ukraine, Kyiv-4, 3, Tereschenkivska st.}
\email{kkuchynskyi@imath.kiev.ua}
\thanks{This work was supported by a grant from the Simons Foundation (1290607, K.O.K.).}

\subjclass[2020]{60H10, 60B05, 60G57}
\keywords{common noise, interacting SDEs, stochastic flows, logarithmic Sobolev inequality, Jacobian matrix flow}

\begin{document}
% TODO: Rewrite abstract
\begin{abstract}
	We study the measure-valued solution to the stochastic differential equations with interaction.
	We derive pathwise transfer estimates for logarithmic Sobolev, Poincar\'e, concentration, and \( T_2 \) inequalities.
	The resulting constants are controlled by the maximal Jacobian norm or the Lipschitz distortion of the flow on the
	support of the initial measure. Under global spatial regularity assumptions,
	we obtain moment estimates for the maximal Jacobian norm on compact sets.
	We also establish an averaged Jacobian-energy bound under one-sided derivative assumptions.
	Finally, for a kernel-type interacting model satisfying a pairwise dissipativity condition,
	we prove exponential decay of the expected pairwise variance, yielding mean-square synchronization around
	the random center of mass.
\end{abstract}

\maketitle

% Navigation: generated table of contents for PDF/source orientation.
\tableofcontents

% ============================================================
% Introduction
% ============================================================
\section{Introduction}\label{sec:introduction}

% The introduction of the main model of the stochastic differential equations with interactions(SDEWI)
We consider stochastic differential equations with interaction that were proposed by Andrey A. Dorogovtsev in \cite{dorogovtsev2003},
%  Main SDEWIN equation
\begin{equation}\label{eq:main-sdewi}
	\begin{cases}
		dX_t(u) & = a\big(X_t(u),\mu_t\big)\,dt + \displaystyle\sum_{\ell=1}^{r} \sigma_\ell\bigl(X_t(u))\,dB_t^\ell, \\[1mm]
		X_0(u)  & = u,                                                                                                \\[1mm]
		\mu_t   & = \mu_0\circ X_t^{-1}.
	\end{cases}
\end{equation}
% The details of the SDEWI
Here \(\mu_0\) is a deterministic initial probability measure, where \(\cP_2(\R^d)\) denotes the space of probability measures
on \(\R^d\) with finite second moment. \(a:\R^d\times\cP_2(\R^d)\to\R^d\) is the drift and the diffusion coefficient
\(\sigma_\ell:\R^d\to\R^d, \quad \ell=1,\ldots,r,\) is state-dependent and non-constant. Let
\( (\Omega,\mathcal F,(\mathcal F_t)_{t\ge0},\mathbb P) \) be a filtered probability space satisfying the usual
conditions and carrying a \(r\)-dimensional Brownian motion \(B=(B^1,\dots,B^r) \).

% Matrix norm
We write \(|\cdot|\) for the Euclidean norm on \(\R^d\). For a matrix \(M\in\R^{d\times d}\) we set
\begin{equation*}
	\|M\|_{\HS}:=\Big(\sum_{i,j=1}^d |M_{ij}|^2\Big)^{1/2}.
\end{equation*}

% Assumption on the coefficients
Fix \(\mu_0\in\mathcal P_2(\mathbb R^d)\).
\begin{assumption}[Well-posedness]\label{ass:wellposedness}
	There exists \(L<\infty\) such that, for all
	\(x,y\in\mathbb R^d\) and
	\(\mu,\nu\in\mathcal P_2(\mathbb R^d)\),
	\begin{equation*}
		|a(x,\mu)-a(y,\nu)| + \left( \sum_{\ell=1}^{r} |\sigma_\ell(x)-\sigma_\ell(y)|^2 \right)^{1/2}  \le L\bigl(|x-y|+W_2(\mu,\nu)\bigr)
	\end{equation*}
\end{assumption}

% The remarks on the solution
Under \Cref{ass:wellposedness}, the SDEWI~\eqref{eq:main-sdewi} admits a unique global solution; see \cite[Theorem~2.3.1]{dorogovtsev2023}. This assumption
remains in force throughout \Cref{sec:jacobian-bound,sec:averaged-jacobian}. Additional spatial regularity conditions are stated locally when they
are needed. \Cref{sec:internal-fluctuations} concerns a different kernel model and is treated under separate assumptions.

% The map of SDEWI as a random transport. The central question of this paper
Unlike a classical stochastic differential equation, the drift depends on the push-forward of the entire system through the random measure \(\mu_t \).
This dependence represents the interaction between particles. For each realization \(\omega \in \Omega\), the map
\begin{equation*}
	T_t(\omega,\cdot):u\mapsto X_t(u,\omega)
\end{equation*}
acts on the initial measure by pushforward,
\begin{equation*}
	\mu_t(\omega)=\mu_0\circ T_t(\omega)^{-1},
\end{equation*}
so the evolution is best viewed as a random transport of measures by a stochastic flow \eqref{eq:main-sdewi}. The central question of this paper is
how the random transport \(T_t\) deforms the properties of the initial measure \( \mu_0 \). Whenever the flow is
differentiable with respect to the initial label, we write
\begin{equation*}
	J_t(u):= D_uT_t(u),
\end{equation*}
and consider the three quantities described below.

% The maximal Jacobian norm
Firstly, we consider the maximal Jacobian norm \(\mathcal K_t(K)\) measures the worst-case infinitesimal stretching of the transport
on the compact set \(K\),
\begin{equation*}
	\mathcal K_t(K):=\sup_{u\in K}\|J_t(u)\|_{\op},
\end{equation*}

% The Jacobian energy 
The Jacobian energy \(\mathcal E_t\)
\begin{equation*}
	\mathcal E_t := \int_{\mathbb R^d}\|J_t(u)\|_{\HS}^2\,\mu_0(du),
\end{equation*}
measures the same infinitesimal deformation after averaging over the initial mass.

% The pairwise energy
The pairwise energy \(\mathcal D_t\)
\begin{equation*}
	\mathcal D_t :=	\iint_{\mathbb R^d\times\mathbb R^d} |T_t(u)-T_t(v)|^2\,\mu_0(du)\mu_0(dv).
\end{equation*}
measures deformation at finite scale by averaging the squared distance between two transported labels.

% Overviiew of these values
These quantities describe different spatial resolutions of the same
random transport. The first is local and unform in the initial label,
the second is local but averaged over the initial distribution, and the
third is averaged and finite-scale. Correspondingly, they lead to
different conclusions and require different assumptions. Uniform Jacobian control is strong enough to transfer functional inequalities
pathwise, but it requires substantial spatial regularity. Averaged Jacobian energy can be estimated under weaker one-sided derivative
conditions. Pairwise energy can be controlled directly through dissipativity of the dynamics, without first obtaining a uniform
Jacobian estimate.

% Section 2 motivation 
Our results follow these three levels of deformation. In \Cref{sec:jacobian-bound}, we study the strongest, worst-case
infinitesimal quantity \(\mathcal K_t(K)\). We prove moment estimates for this random distortion constant on compact sets and show that it
governs the pathwise transfer of logarithmic Sobolev, Poincar\'e, concentration, and transport inequalities from \(\mu_0\) to \(\mu_t\).

% Section 3 motivation
In \Cref{sec:averaged-jacobian}, we lower the level of spatial resolution and replace the supremum over labels by the averaged
Jacobian energy \(\mathcal E_t\). This quantity does not provide the same uniform pathwise information, but it admits a direct estimate
under weaker one-sided assumptions on the derivative coefficients.

% Section 4 summary
Finally, in \Cref{sec:internal-fluctuations}, we pass from infinitesimal deformation to finite-scale deformation. For a
kernel-type interacting model, we impose a pairwise dissipativity condition and prove exponential decay of \(\mathbb E\mathcal D_t\).
This decay means that the transported labels approach one another in mean square and that the random measure synchronizes around its random
center of mass.

\clearpage

% ============================================================
% Pathwise Jacobian Bounds and Functional Inequality Transfer
% ============================================================
\section{Pathwise Jacobian Bounds and Functional Inequality Transfer}\label{sec:jacobian-bound}

% Roadmap to the section
This section has two purposes. First, we quantify the spatial distortion of the stochastic flow \( T_t \) through its Jacobian \(J_t \). Second,
we use this distortion to transfer functional inequalities from the initial measure \( \mu_0 \) to the random pushforward measure \( \mu_t \). The
technical proof of the compact-set Jacobian estimate is postponed to the final subsection.

\begin{assumption}[Spatial differentiability]\label{ass:flow-differentiability}
	For every \(\mu\in\mathcal P_2(\mathbb R^d)\), the map \(x\mapsto a(x,\mu)\) belongs to
	\(C^1(\mathbb R^d;\mathbb R^d)\), and
	\begin{equation*}
		\sigma_\ell\in C^1(\mathbb R^d;\mathbb R^d), \qquad \ell=1,\dots,r.
	\end{equation*}
\end{assumption}

By \Cref{ass:wellposedness,ass:flow-differentiability},
\begin{equation}\label{eq:basic-derivative-bounds}
	\|D_xa(x,\mu)\|_{\op}\le L_x, \qquad \sum_{\ell=1}^{r}|D\sigma_\ell(x)z|^2 \le L_\sigma^2|z|^2
\end{equation}
for all \(x,z\in\mathbb R^d\) and \(\mu\in\mathcal P_2(\mathbb R^d)\)

% ============================================================
% The Jacobian flow and distortion constants
% ============================================================
\subsection{The Jacobian flow and distortion constants}\label{subsec:jacobian-flow-distortion-constants}

% Roadmap to the subsection
We first identify the Jacobian equation for the stochastic flow and introduce the distortion quantities used throughout the section.

For a fixed \(\omega\) and \(t\), define the flow map \(T_t(\omega,u):=X_t(u,\omega)\) and its Jacobian matrix
\begin{equation*}
	J_t(u):=\partial_u T_t(u)=\partial_u X_t(u)\in\R^{d\times d}.
\end{equation*}
We use the Jacobian SDE from \cite{dorogovtsev_weiss2023}.
For \cref{eq:main-sdewi}, it has the following form.

\begin{theorem}[Jacobian equation] \label{thm:jacobian-sde}
	Assume \Cref{ass:wellposedness,ass:flow-differentiability}. Then the solution admits a
	modification such that, for almost every \(\omega\) and every \(t\ge0\), the map
	\begin{equation*}
		u\longmapsto T_t(\omega,u)
	\end{equation*}
	is of class \(C^1\). Its Jacobian \( J_t(u) \) satisfies
	\begin{equation}\label{eq:jacobian-sde}
		\begin{cases}
			dJ_t(u) = A_t(u)J_t(u)\,dt + \sum_{\ell=1}^{r} C_{\ell,t}(u)J_t(u)\,dB_t^\ell, \\
			J_0(u)=I,                                                                      \\
		\end{cases}
	\end{equation}
	where
	\begin{equation*}
		A_t(u):=D_xa\bigl(X_t(u),\mu_t\bigr),
		\qquad
		C_{\ell,t}(u):=D\sigma_\ell\bigl(X_t(u)\bigr).
	\end{equation*}
	In particular,
	\begin{equation*}
		\|A_t(u)\|_{\op}\le L_x, \qquad \sum_{\ell=1}^{r} |C_{\ell,t}(u)z|^2 \le L_\sigma^2|z|^2
	\end{equation*}
	for all \(t,u,z\).
\end{theorem}

\begin{definition}[Random Jacobian constant]\label{def:jacobian-constant}
	For a nonempty set \(K\subset\R^d\), define
	\begin{equation*}
		\mathcal{K}_t(\omega;K):=\sup_{u\in K}\|J_t(u,\omega)\|_{\op}.
	\end{equation*}
	If \(K\) is compact and \(u\mapsto J_t(u,\omega)\) is continuous, then \(\mathcal{K}_t(\omega;K)<\infty\).
\end{definition}

% Remark
If \(K\) is compact, continuity of \(u\mapsto J_t(u)\) implies \(\mathcal K_t(K)<\infty\). This compact-set quantity is used because a
corresponding supremum over all of \(\mathbb R^d\) need not be finite. The counterexamples of Mohammed--Scheutzow show that even for globally
Lipschitz coefficients one may fail to have a finite global constant \(\sup_{u\in\R^d}\|J_t(u,\omega)\|_{\op}\) for a fixed \((t,\omega)\).

\begin{lemma}[Flow Lipschitz constant via Jacobian matrix]\label{lem:lip-via-jacobian}
	For each fixed \((t,\omega)\) such that the flow map \(u\mapsto T_t(\omega,u):=X_t(u,\omega)\) is of class \(C^1\),
	\begin{equation*}
		\Lip(T_t(\omega)) = \sup_{u\in\R^d}\|J_t(u,\omega)\|_{\op}\in[0,\infty].
	\end{equation*}
\end{lemma}

\begin{proof}
	Since \(u\mapsto T_t(\omega,u)\) is \(C^1\), the mean value theorem implies
	\begin{equation*}
		|T_t(u)-T_t(v)| \le \sup_{\theta\in[0,1]}\|J_t(v+\theta(u-v))\|_{\op}\,|u-v|.
	\end{equation*}
	Taking the supremum over \(u\neq v\) yields \(\Lip(T_t)\le \sup_u\|J_t(u)\|_{\op}\).
	Conversely, for any \(u\) and unit vector \(\xi\), letting \(v=u+\varepsilon\xi\) and sending \(\varepsilon\downarrow 0\) gives
	\begin{equation*}
		\|J_t(u)\|_{\op} = \sup_{|\xi|=1}\lim_{\varepsilon\downarrow 0}\frac{|T_t(u+\varepsilon\xi)-T_t(u)|}{\varepsilon}
		\le \Lip(T_t),
	\end{equation*}
	hence equality.
\end{proof}

% Pointwise estimates control \(\|J_t(u)\|_{\op}\) for fixed \(u\). To control the random Lipschitz constant
% \begin{equation*}
%     \mathcal{K}_t(\omega;K):=\sup_{u\in K}\|J_t(u,\omega)\|_{\op}
% \end{equation*}
% on a compact set \(K\), we impose the following additional spatial regularity assumption.
% \begin{assumption}\label{ass:spatial-lip-derivatives}
%     There exist constants \(L^{(2)}<\infty\) and \(L_\sigma^{(2)}<\infty\) such that for all \(\mu\in\cP_2(\R^d)\) and all \(x,y\in\R^d\),
%     \begin{align*}
%         \|\partial_x a(x,\mu)-\partial_x a(y,\mu)\|_{\op}                             & \le L^{(2)}\,|x-y|,        \\
%         \max_{1\le k\le d}\|\partial \sigma^{(k)}(x)-\partial \sigma^{(k)}(y)\|_{\op} & \le L_\sigma^{(2)}\,|x-y|.
%     \end{align*}
% \end{assumption}

% \begin{theorem}[Moment bound for \(\mathcal{K}_t(\cdot;K)\)]\label{thm:kt-moment}
%     Assume \Cref{ass:general-lipschitz,ass:spatial-lip-derivatives}. Then for every compact \(K\subset\R^d\), every \(p\ge 1\), and every \(t\ge0\),
%     \begin{equation}\label{eq:kt-finite-moment}
%         \E\,\mathcal{K}_t(\cdot;K)^p<\infty.
%     \end{equation}
%     Moreover, there exist constants \(C_{p,K},c_p<\infty\) (depending only on \(p,K,d,L,L_\sigma,L^{(2)},L_\sigma^{(2)}\)) such that
%     \begin{equation}\label{eq:kt-exp-growth}
%         \E\,\mathcal{K}_t(\cdot;K)^p \le C_{p,K}\,e^{c_p t}, \qquad t\ge0.
%     \end{equation}
% \end{theorem}
Pointwise estimates for \(J_t(u)\) follow from the boundedness of the coefficients in the linear equation \eqref{eq:jacobian-sde}. Controlling the
compact-set maximal Jacobian norm \(\mathcal K_t(K)\), however, requires estimates for the spatial increments \(J_t(u)-J_t(v)\). The additional
regularity needed for this purpose is therefore imposed only in the following theorem.

\begin{theorem}[Moments of the compact-set maximal Jacobian norm]\label{thm:kt-moment}
	Assume \Cref{ass:wellposedness,ass:flow-differentiability}. Suppose in addition that there exists a constant
	\(H<\infty\) such that, for all \(x,y,z\in\mathbb R^d\) and \(\mu\in\mathcal P_2(\mathbb R^d)\),
	\begin{equation}\label{eq:spatial-lipschitz-derivatives}
		\big| \big(D_xa(x,\mu)-D_xa(y,\mu)\big)z \big| + \left( \sum_{\ell=1}^{r} \big| \big(D\sigma_\ell(x)-D\sigma_\ell(y)\big)z \big|^2 \right)^{1/2}
		\le H|x-y|\,|z|.
	\end{equation}
	Then, for every nonempty compact \(K\subset\mathbb R^d\) and every
	\(p\ge1\), there exist constants \(C_{p,K},c_p<\infty\) such that
	\begin{equation}\label{eq:kt-exp-growth}
		\mathbb E\big[\mathcal K_t(\cdot;K)^p\big] \le C_{p,K}e^{c_pt}, \qquad t\ge0.
	\end{equation}
\end{theorem}

The proof is given in \Cref{subsec:kt-moment-proof}.

% ============================================================
% Transfer of functional inequalities
% ============================================================
\subsection{Transfer of functional inequalities}\label{subsec:functional-inequality-transfer}

In this subsection we record several consequences of the identity
\begin{equation*}
	T_t(\omega,u):=X_t(u,\omega),
\end{equation*}
which expresses the random measure \(\mu_t\) as the pushforward of the initial law \(\mu_0\) under the stochastic flow.

% Logarithmic Sobolev inequality
\begin{definition}[Logarithmic Sobolev inequality]
	Denoting by \(\mathcal{C}^1_c(\R^d)\) the set of continuously differentiable compactly supported functions from \(\R^d\) to \(\R\), for \(g\ge 0\) with \(\int g\,d\mu=1\) we define the entropy
	\begin{equation*}
		\Ent_\mu(g):=\int g\log g\,d\mu.
	\end{equation*}
	A probability measure \(\mu\) on \(\R^d\) is said to satisfy a logarithmic Sobolev inequality with constant \(C>0\) if for every \(h\in\mathcal{C}^1_c(\R^d)\) with \(\int h^2\,d\mu=1\) one has
	\begin{equation}\label{eq:lsi-def}
		\Ent_\mu(h^2)\le C\int_{\R^d}|\nabla h|^2\,d\mu.
	\end{equation}
\end{definition}

\begin{corollary}[Pathwise LSI with the Jacobian matrix constant]\label{cor:pathwise-lsi}
	If \(\mu_0\) satisfies \(\mathrm{LSI}(C_0)\) and \(\supp(\mu_0)\subset K\) for some compact \(K\), then for every \(t\ge0\) and a.s.\ \(\omega\),
	\begin{equation*}
		\mu_t(\omega)=\mu_0\circ T_t(\omega)^{-1}
		\quad\Longrightarrow\quad
		\mu_t(\omega)\text{ satisfies }\mathrm{LSI}\big(C_0\,\mathcal{K}_t(\omega;K)^2\big).
	\end{equation*}
\end{corollary}

We next record the basic concentration-transfer principle under a random Lipschitz map.

\begin{proposition}[Pathwise concentration transfer]\label{prop:concentration-transfer}
	Let \(K:=\supp(\mu_0)\) and assume that for a.s.\ \(\omega\) the map \(T_t(\omega,\cdot)\) is Lipschitz on \(K\), with Lipschitz constant
	\begin{equation*}
		L_t(\omega;K):=\sup_{\substack{u,v\in K\\u\neq v}}
		\frac{|T_t(\omega,u)-T_t(\omega,v)|}{|u-v|}.
	\end{equation*}
	Assume that \(\mu_0\) satisfies the concentration inequality
	\begin{equation}\label{eq:general-concentration-mu0}
		\mu_0(f-\mu_0(f)\ge r)\le \Psi(r),
		\qquad r\ge0,
	\end{equation}
	for every \(1\)-Lipschitz function \(f:\R^d\to\R\), where \(\Psi:[0,\infty)\to[0,1]\) is nonincreasing.
	Then for a.s.\ \(\omega\),
	\begin{equation}\label{eq:general-concentration-mut}
		\mu_t(\omega)(g-\mu_t(\omega)(g)\ge r)
		\le
		\Psi\Big(\frac{r}{L_t(\omega;K)}\Big),
		\qquad r\ge0,
	\end{equation}
	for every \(1\)-Lipschitz function \(g:\R^d\to\R\).
\end{proposition}

\begin{proof}
	Fix \(\omega\) such that \(T_t(\omega,\cdot)\) is Lipschitz on \(K\), and write
	\(L:=L_t(\omega;K)\in[0,\infty)\).
	Let \(g:\R^d\to\R\) be \(1\)-Lipschitz.

	If \(L=0\), then \(T_t(\omega,\cdot)\) is constant on \(K\), hence \(\mu_t(\omega)=\mu_0\circ T_t(\omega)^{-1}\) is a Dirac mass and
	\(\mu_t(\omega)(g-\mu_t(\omega)(g)\ge r)=0\) for every \(r>0\). Thus \cref{eq:general-concentration-mut} is immediate.

	Assume now \(L>0\) and define on \(K\)
	\begin{equation*}
		h(u):=\frac{g(T_t(\omega,u))}{L}.
	\end{equation*}
	Then \(h\) is \(1\)-Lipschitz on \(K\).
	By the McShane extension theorem, \(h\) extends to a \(1\)-Lipschitz function
	\(\widetilde h:\R^d\to\R\).
	Applying \cref{eq:general-concentration-mu0} to \(\widetilde h\) (and using that \(\mu_0\) is supported on \(K\)) gives
	\begin{equation*}
		\mu_0\big(h-\mu_0(h)\ge s\big)\le \Psi(s),\qquad s\ge0.
	\end{equation*}
	Since \(\mu_t(\omega)=\mu_0\circ T_t(\omega)^{-1}\),
	\begin{equation*}
		\mu_t(\omega)(g)=\mu_0(g\circ T_t(\omega,\cdot))=L\mu_0(h).
	\end{equation*}
	Therefore, for \(r\ge0\),
	\begin{align*}
		\mu_t(\omega)(g-\mu_t(\omega)(g)\ge r)
		 & = \mu_0\big(g\circ T_t(\omega,\cdot)-\mu_0(g\circ T_t(\omega,\cdot))\ge r\big) \\
		 & = \mu_0\big(h-\mu_0(h)\ge r/L\big)                                             \\
		 & \le \Psi\Big(\frac{r}{L}\Big)
		= \Psi\Big(\frac{r}{L_t(\omega;K)}\Big).
	\end{align*}
	This proves \cref{eq:general-concentration-mut}.
\end{proof}

\begin{remark}[Concentration as a corollary of LSI]
	If \(\mu_0\) satisfies a logarithmic Sobolev inequality, then it satisfies Gaussian concentration.
	Hence \Cref{cor:pathwise-lsi} yields, in particular, a pathwise concentration inequality for \(\mu_t(\omega)\) with a random constant controlled by \(\mathcal K_t(\omega;K)^2\).
\end{remark}

% Poincare inequality
\begin{definition}[Poincar\'e inequality]
	A probability measure \(\mu\) on \(\R^d\) is said to satisfy a Poincar\'e inequality with constant \(C_P>0\), denoted by \(\mathrm{PI}(C_P)\), if for every \(h\in \mathcal C_c^1(\R^d)\),
	\begin{equation}\label{eq:poincare-def}
		\Var_\mu(h):=\int_{\R^d} h^2\,d\mu-\Big(\int_{\R^d} h\,d\mu\Big)^2
		\le
		C_P\int_{\R^d} |\nabla h|^2\,d\mu.
	\end{equation}
\end{definition}

We first record the stability of the Poincar\'e inequality under Lipschitz pushforward.

\begin{proposition}[Poincar\'e inequality under a Lipschitz pushforward]\label{prop:poincare-pushforward}
	Let \(\mu\) be a probability measure on \(\R^d\) satisfying \(\mathrm{PI}(C_P)\), and let
	\(T:\R^d\to\R^d\) be a Lipschitz map with Lipschitz constant \(\Lip(T)\).
	Then the pushforward measure \(\nu:=\mu\circ T^{-1}\) satisfies
	\begin{equation*}
		\mathrm{PI}\big(C_P\,\Lip(T)^2\big).
	\end{equation*}
\end{proposition}

\begin{proof}
	Let \(g\in \mathcal C_c^1(\R^d)\) and set \(f:=g\circ T\).
	Then
	\begin{equation*}
		\int_{\R^d} g\,d\nu=\int_{\R^d} g(T(u))\,\mu(du)=\int_{\R^d} f(u)\,\mu(du),
	\end{equation*}
	and similarly
	\begin{equation*}
		\int_{\R^d} g^2\,d\nu=\int_{\R^d} f^2\,d\mu.
	\end{equation*}
	Hence
	\begin{equation*}
		\Var_\nu(g)=\Var_\mu(f).
	\end{equation*}
	Applying the Poincar\'e inequality for \(\mu\) gives
	\begin{equation*}
		\Var_\nu(g) = \Var_\mu(f) \le C_P\int_{\R^d} |\nabla f(u)|^2\,\mu(du).
	\end{equation*}
	Since \(f=g\circ T\) and \(T\) is Lipschitz, we have a.e.
	\begin{equation*}
		|\nabla f(u)| \le \Lip(T)\,|\nabla g(T(u))|.
	\end{equation*}
	Therefore
	\begin{equation*}
		\Var_\nu(g) \le C_P\,\Lip(T)^2 \int_{\R^d} |\nabla g(T(u))|^2\,\mu(du) = C_P\,\Lip(T)^2 \int_{\R^d} |\nabla g(x)|^2\,\nu(dx).
	\end{equation*}
	This proves the claim.
\end{proof}

\begin{corollary}[Pathwise Poincar\'e inequality]\label{cor:pathwise-poincare}
	If \(\mu_0\) satisfies \(\mathrm{PI}(C_P)\) and \(\supp(\mu_0)\subset K\) for some compact set \(K\subset \R^d\), then for every \(t\ge0\) and a.s.\ \(\omega\),
	\begin{equation*}
		\mu_t(\omega)=\mu_0\circ T_t(\omega)^{-1}
		\quad\Longrightarrow\quad
		\mu_t(\omega)\text{ satisfies }\mathrm{PI}\big(C_P\,\mathcal K_t(\omega;K)^2\big).
	\end{equation*}
\end{corollary}

\begin{proof}
	By \Cref{prop:poincare-pushforward}, it suffices to bound the Lipschitz constant of \(T_t(\omega,\cdot)\) on \(K\).
	Since \(K\) is compact and \(u\mapsto T_t(\omega,u)\) is of class \(C^1\), the mean value theorem gives
	\begin{equation*}
		|T_t(\omega,u)-T_t(\omega,v)| \le \sup_{z\in K}\|J_t(z,\omega)\|_{\op}\,|u-v| = \mathcal K_t(\omega;K)\,|u-v|, \qquad u,v\in K.
	\end{equation*}
	Hence \(\Lip(T_t(\omega)|_K)\le \mathcal K_t(\omega;K)\), and the result follows.
\end{proof}

% Talagrand transport inequality
\begin{definition}[Talagrand transport inequality]
	Let \(p\ge1\). A probability measure \(\mu\) on \(\R^d\) is said to satisfy the Talagrand transport inequality \(T_p(C_T)\) with constant \(C_T>0\) if for every probability measure \(\nu\) on \(\R^d\) with finite \(p\)-th moment,
	\begin{equation}\label{eq:talagrand-def}
		W_p(\nu,\mu)^2 \le 2C_T\,\Ent(\nu\mid \mu),
	\end{equation}
	where
	\begin{equation*}
		\Ent(\nu\mid\mu)=
		\begin{cases}
			\displaystyle \int_{\R^d}\log\Big(\frac{d\nu}{d\mu}\Big)\,d\nu, & \nu\ll \mu,       \\[3mm]
			+\infty,                                                        & \text{otherwise}.
		\end{cases}
	\end{equation*}
\end{definition}

We next record the analogous stability property for Talagrand transport inequalities.

\begin{proposition}[Transport inequality under a Lipschitz pushforward]\label{prop:talagrand-pushforward}
	Let \(\mu\) be a probability measure on \(\R^d\) satisfying \(T_p(C_T)\) for some \(p\ge1\), and let
	\(T:\R^d\to\R^d\) be a Lipschitz map with Lipschitz constant \(\Lip(T)\).
	Then the pushforward measure \(\eta:=\mu\circ T^{-1}\) satisfies
	\begin{equation*}
		T_p\big(C_T\,\Lip(T)^2\big).
	\end{equation*}
\end{proposition}

\begin{proof}
	Let \(\lambda\) be any probability measure on \(\R^d\).
	If \(\lambda\not\ll \eta\), then \(\Ent(\lambda\mid\eta)=+\infty\) and the inequality is trivial.
	Assume therefore that \(\lambda\ll \eta\).

	Choose a probability measure \(\widetilde\lambda\) on \(\R^d\) such that
	\begin{equation*}
		\widetilde\lambda\circ T^{-1}=\lambda
		\qquad\text{and}\qquad
		\Ent(\widetilde\lambda\mid \mu)=\Ent(\lambda\mid \eta).
	\end{equation*}
	For instance, one may take the measure
	\begin{equation*}
		\widetilde\lambda(du):=
		\frac{d\lambda}{d\eta}(T(u))\,\mu(du).
	\end{equation*}
	Indeed,
	\begin{equation*}
		(\widetilde\lambda\circ T^{-1})(A)
		=
		\int_{\R^d}\mathbf 1_{\{T(u)\in A\}}\frac{d\lambda}{d\eta}(T(u))\,\mu(du)
		=
		\int_A \frac{d\lambda}{d\eta}(x)\,\eta(dx)
		=
		\lambda(A),
	\end{equation*}
	and
	\begin{equation*}
		\Ent(\widetilde\lambda\mid\mu)
		=
		\int_{\R^d}\log\Big(\frac{d\lambda}{d\eta}(T(u))\Big)\,\widetilde\lambda(du)
		=
		\int_{\R^d}\log\Big(\frac{d\lambda}{d\eta}(x)\Big)\,\lambda(dx)
		=
		\Ent(\lambda\mid\eta).
	\end{equation*}

	Now let \(\pi\) be a coupling of \(\widetilde\lambda\) and \(\mu\). Then \(\pi\circ (T,T)^{-1}\) is a coupling of \(\lambda\) and \(\eta\), so by
	Lipschitz continuity of \(T\),
	\begin{align*}
		 & \int_{\R^d\times\R^d}
		|x-y|^p\,d(\pi\circ (T,T)^{-1})(x,y)          \\
		 & \quad =
		\int_{\R^d\times\R^d}|T(u)-T(v)|^p\,d\pi(u,v) \\
		 & \quad \le
		\Lip(T)^p
		\int |u-v|^p\,d\pi(u,v).
	\end{align*}
	Taking the infimum over all couplings \(\pi\) of \(\widetilde\lambda\) and \(\mu\) yields
	\begin{equation*}
		W_p(\lambda,\eta)\le \Lip(T)\,W_p(\widetilde\lambda,\mu).
	\end{equation*}
	Since \(\mu\) satisfies \(T_p(C_T)\),
	\begin{equation*}
		W_p(\widetilde\lambda,\mu)^2
		\le
		2C_T\,\Ent(\widetilde\lambda\mid\mu)
		=
		2C_T\,\Ent(\lambda\mid\eta).
	\end{equation*}
	Combining the two inequalities, we obtain
	\begin{equation*}
		W_p(\lambda,\eta)^2
		\le
		\Lip(T)^2\,W_p(\widetilde\lambda,\mu)^2
		\le
		2C_T\,\Lip(T)^2\,\Ent(\lambda\mid\eta),
	\end{equation*}
	which proves that \(\eta\) satisfies \(T_p(C_T\,\Lip(T)^2)\).
\end{proof}

% ============================================================
% Proof of the compact-set Jacobian estimate
% ============================================================
\subsection{Proof of the compact-set Jacobian estimate}\label{subsec:kt-moment-proof}

The proof of \Cref{thm:kt-moment} relies on two elementary estimates for linear stochastic equations. The first controls moments of a homogeneous
linear equation and will be applied both to differences of flow trajectories and to the columns of the Jacobian matrix. The second treats an
inhomogeneous equation and will be used to estimate spatial increments of the Jacobian. We state these estimates separately in order to keep the
proof of the theorem focused on the spatial structure of the flow.

\begin{lemma}[Moment estimate for linear stochastic equations]\label{lem:linear-sde-moment}
	Let \(Y_t\) solve
	\begin{equation*}
		\begin{cases}
			dY_t = A_tY_t\,dt + \sum_{\ell=1}^{r} C_{\ell,t}Y_t\,dB_t^\ell, \\
			Y_0=y.                                                          \\
		\end{cases}
	\end{equation*}

	Assume that, for every \(z\in\R^d\),
	\begin{equation*}
		\langle z,A_tz\rangle \le L_x|z|^2, \qquad
		\sum_{\ell=1}^{r}|C_{\ell,t}z|^2  \le L_\sigma^2|z|^2.
	\end{equation*}

	Then, for every \(q\ge1\),
	\begin{equation*}
		\E|Y_t|^{2q} \le |y|^{2q} \exp\left\{ q\bigl(2L_x+(2q-1)L_\sigma^2\bigr)t \right\}.
	\end{equation*}

	Consequently,
	\begin{equation*}
		\E\|J_t(u)\|_{\op}^{2q} \le d^q \exp\left\{ q\bigl(2L_x+(2q-1)L_\sigma^2\bigr)t \right\}.
	\end{equation*}
\end{lemma}

\noindent\emph{Proof.}
Applying It\^o's formula to \(|Y_t|^{2q}\), after a standard
localization, gives
\begin{align*}
	d|Y_t|^{2q} & = 2q|Y_t|^{2q-2}\langle Y_t,A_tY_t\rangle\,dt                                              \\
	            & \quad+q|Y_t|^{2q-2} \sum_{\ell=1}^{r}|C_{\ell,t}Y_t|^2\,dt                                 \\
	            & \quad +2q(q-1)|Y_t|^{2q-4} \sum_{\ell=1}^{r} \langle Y_t,C_{\ell,t}Y_t\rangle^2\,dt +dM_t,
\end{align*}
where \(M_t\) is a local martingale. Since
\begin{equation*}
	\langle Y_t,C_{\ell,t}Y_t\rangle^2 \le |Y_t|^2|C_{\ell,t}Y_t|^2,
\end{equation*}
the assumptions imply
\begin{equation*}
	d|Y_t|^{2q} \le q\bigl(2L_x+(2q-1)L_\sigma^2\bigr) |Y_t|^{2q}\,dt+dM_t.
\end{equation*}
Taking expectations in the localized equation and applying Gronwall's
lemma yields
\begin{equation*}
	\E|Y_t|^{2q} \le |y|^{2q} \exp\left\{ q\bigl(2L_x+(2q-1)L_\sigma^2\bigr)t \right\}.
\end{equation*}

For the matrix estimate, apply the vector estimate to \(J_t(u)e_j\), \(j=1,\dots,d\), and use
\begin{equation*}
	\|J_t(u)\|_{\op}^{2q} \le \|J_t(u)\|_{\HS}^{2q} = \left(\sum_{j=1}^{d}|J_t(u)e_j|^2\right)^q
	\le d^{q-1}\sum_{j=1}^{d}|J_t(u)e_j|^{2q}.
\end{equation*}
This gives
\begin{equation*}
	\E\|J_t(u)\|_{\op}^{2q} \le d^q \exp\left\{ q\bigl(2L_x+(2q-1)L_\sigma^2\bigr)t \right\}.
\end{equation*}
\hfill\(\Box\)

The preceding lemma applies directly to \(X_t(u)-X_t(v)\), because the difference of two trajectories driven by the same Brownian motion satisfies
a homogeneous linear equation with random coefficients. The difference \(J_t(u)-J_t(v)\), however, satisfies an inhomogeneous linear equation: the
additional terms arise from evaluating the derivative coefficients along the two different trajectories. We therefore record the corresponding
moment estimate.

\begin{lemma}[Inhomogeneous linear moment estimate]
	\label{lem:inhomogeneous-linear-moment}
	Let \(m\ge2\) be an even integer, and suppose that the
	matrix-valued process \(D_t\) satisfies
	\begin{equation*}
		dD_t = A_tD_t\,dt +\sum_{\ell=1}^{r}C_{\ell,t}D_t\,dB_t^\ell
		+F_t\,dt +\sum_{\ell=1}^{r}G_{\ell,t}\,dB_t^\ell.
	\end{equation*}
	Assume that, for every \(M\in\mathbb R^{d\times d}\),
	\begin{equation*}
		\langle M,A_tM\rangle_{\HS} \le L_x\|M\|_{\HS}^2, \qquad
		\sum_{\ell=1}^{r}\|C_{\ell,t}M\|_{\HS}^2 \le L_\sigma^2\|M\|_{\HS}^2.
	\end{equation*}
	Then there exist constants \(C_m,c_m<\infty\), depending only on
	\(m,L_x,L_\sigma\), such that
	\begin{align*}
		\E\|D_t\|_{\HS}^m & \le e^{c_mt}\E\|D_0\|_{\HS}^m                            \\
		                  & \quad +C_m\int_0^t e^{c_m(t-s)} \left[ \E\|F_s\|_{\HS}^m
		+ \E\left( \sum_{\ell=1}^{r}\|G_{\ell,s}\|_{\HS}^2 \right)^{m/2} \right]ds.
	\end{align*}
\end{lemma}

\noindent\emph{Proof.}
Apply It\^o's formula to \(\|D_t\|_{\HS}^m\), after localization.
Using
\begin{equation*}
	\sum_{\ell=1}^{r}
	\|C_{\ell,t}D_t+G_{\ell,t}\|_{\HS}^2
	\le
	2L_\sigma^2\|D_t\|_{\HS}^2
	+
	2\sum_{\ell=1}^{r}\|G_{\ell,t}\|_{\HS}^2
\end{equation*}
and Young's inequality gives
\begin{align*}
	\frac{d}{dt}\E\|D_t\|_{\HS}^m
	 & \le
	c_m\E\|D_t\|_{\HS}^m
	+C_m\E\|F_t\|_{\HS}^m \\
	 & \quad
	+C_m\E\left(
	\sum_{\ell=1}^{r}\|G_{\ell,t}\|_{\HS}^2
	\right)^{m/2}.
\end{align*}
The result follows from Gronwall's lemma.
\hfill\(\Box\)

Although the drift depends on the random measure \( \mu_t \), differentiation with respect to the initial label acts only on the spatial argument
\(X_t(u) \). Because of that fact we can evaluate \(\|J_t(u)\|\) using Kunita's theorem for the stochastic flows.

Under \Cref{ass:wellposedness,ass:flow-differentiability}, standard results on stochastic flows (see, e.g., \cite[Theorem~1.4.1]{Kunita1986})
yield a modification such that for a.s.\ \(\omega\) and each \(t\ge0\) the map \(u\mapsto T_t(\omega,u)\) is \(\mathcal{C}^1\).

\begin{proof}[Proof of \Cref{thm:kt-moment}]

	% fix initial conditions
	Fix a compact set \(K\subset\R^d\), \(t\ge0\), and \(p\ge1\). Choose an even integer \(m>\max\{2p,2d\}\). It is enough to prove
	\eqref{eq:kt-exp-growth} with \(p\) replaced by \(m\), because then H\"older's inequality gives
	\begin{equation*}
		\E\,\mathcal{K}_t(\cdot;K)^p \le \big(\E\,\mathcal{K}_t(\cdot;K)^m\big)^{p/m}.
	\end{equation*}

	% define the difference \Delta X_t
	Fix \(u,v\in\R^d\) and set
	\begin{equation*}
		\Delta X_t:=X_t(u)-X_t(v).
	\end{equation*}

	% explicit d\Delta X_t
	Since the same random measure \(\mu_t\) appears in both equations, subtraction yields
	\begin{equation*}
		d\Delta X_t = \bigl( a(X_t(u),\mu_t)-a(X_t(v),\mu_t) \bigr)\,dt + \sum_{\ell=1}^{r} \bigl(\sigma_\ell(X_t(u))-\sigma_\ell(X_t(v)) \bigr)\,dB_t^\ell.
	\end{equation*}

	% the average coefficients \widetilde A_t(u,v) and \widetilde C_{\ell,t}(u,v)
	Define
	\begin{equation*}
		\widetilde A_t(u,v) := \int_0^1 D_xa\bigl( X_t(v)+\theta\Delta X_t,\mu_t \bigr)\,d\theta
	\end{equation*}
	and, for \(\ell=1,\dots,r\),
	\begin{equation*}
		\widetilde C_{\ell,t}(u,v) := \int_0^1 D\sigma_\ell\bigl( X_t(v)+\theta\Delta X_t \bigr)\,d\theta.
	\end{equation*}

	% the sde via fundamental theorem of calculus
	By the fundamental theorem of calculus,
	\begin{equation*}
		\begin{cases}
			d\Delta X_t = \widetilde A_t(u,v)\Delta X_t\,dt + \sum_{\ell=1}^{r} \widetilde C_{\ell,t}(u,v)\Delta X_t\,dB_t^\ell, \\[2mm]
			\Delta X_0=u-v.                                                                                                      \\[2mm]
		\end{cases}
	\end{equation*}

	% making sure the condition of lemma hold true
	Moreover, by \eqref{eq:basic-derivative-bounds},
	\begin{equation*}
		\|\widetilde A_t(u,v)\|_{\op}\le L_x.
	\end{equation*}
	Jensen's inequality also gives, for every \(z\in\mathbb R^d\),
	\begin{equation*}
		\sum_{\ell=1}^{r} |\widetilde C_{\ell,t}(u,v)z|^2
		\le \int_0^1 \sum_{\ell=1}^{r} \left| D\sigma_\ell\bigl( X_t(v)+\theta\Delta X_t \bigr)z \right|^2\,d\theta
		\le L_\sigma^2|z|^2.
	\end{equation*}

	% Applying lemma for the liner sde moments
	Applying \Cref{lem:linear-sde-moment} with \(q=m/2\) and with \(q=m\), respectively, yields
	\begin{align}
		\E|X_t(u)-X_t(v)|^m    & \le C_me^{c_mt}|u-v|^m, \label{eq:flow-increment-moment}       \\
		\E|X_t(u)-X_t(v)|^{2m} & \le C_me^{c_mt}|u-v|^{2m}. \label{eq:flow-increment-moment-2m}
	\end{align}

	% If \(u=v\), then \(\Delta X_t\equiv0\), so the desired estimate is trivial. Assume therefore that \(u\neq v\), and set
	% \begin{equation*}
	%     Z_t:=\frac{\Delta X_t}{|u-v|},
	%     \qquad
	%     \xi:=\frac{u-v}{|u-v|}.
	% \end{equation*}
	% Then \(|\xi|=1\), \(Z_0=\xi\), and \(Z_t\) solves
	% \begin{equation*}
	%     \begin{cases}
	%         dZ_t = \widetilde A_t(u,v)\,Z_t\,dt + \sum_{k=1}^d \widetilde C_t^{(k)}(u,v)\,Z_t\,dB_t^k, \\[2mm]
	%         Z_0 = \frac{u-v}{|u-v|}.                                                                   \\[2mm]
	%     \end{cases}
	% \end{equation*}
	% Applying \Cref{lem:linear-sde-moment} to this linear equation with exponent \(q=m/2\), we obtain
	% \begin{equation*}
	%     \E|Z_t|^m
	%     \le
	%     C_m e^{c_m t},
	% \end{equation*}
	% for constants \(C_m,c_m<\infty\) depending only on \(m,d,L,L_\sigma\).
	% Multiplying by \(|u-v|^m\) yields
	% \begin{equation}\label{eq:flow-increment-moment}
	%     \E\,|X_t(u)-X_t(v)|^m
	%     \le
	%     C_m e^{c_m t}|u-v|^m,
	%     \qquad u,v\in\R^d.
	% \end{equation}

	\Cref{lem:linear-sde-moment} gives, for every \(q\ge1\) and every \(u\in\R^d\),
	\begin{equation*}
		\E\,\|J_t(u)\|_{\op}^{2q}
		\le d^q \exp\Big(q(2L+dL_\sigma^2)t+2q^2 dL_\sigma^2 t\Big).
	\end{equation*}
	In particular, taking \(q=m/2\) and \(q=m\), we obtain
	\begin{equation}\label{eq:jacobian-moment}
		\E\,\|J_t(u)\|_{\op}^m \le C_m e^{c_m t}, \qquad
		\E\,\|J_t(u)\|_{\op}^{2m}
		\le C_m e^{c_m t},
	\end{equation}
	uniformly in \(u\in\R^d\).

	\paragraph{Step 2: Spatial increments of the Jacobian.}

	% define the difference D_t
	Fix \(u,v\in\mathbb R^d\), and write
	\begin{equation*}
		D_t:=J_t(u)-J_t(v).
	\end{equation*}
	% define the difference D_t
	Set
	\begin{equation*}
		A_t(w):=D_x a(X_t(w),\mu_t), \qquad
		C_{\ell,t}(w):=D\sigma_\ell(X_t(w)), \qquad \ell=1,\dots,r.
	\end{equation*}
	% set the sde for D_t equation
	Subtracting the Jacobian equations at \(u\) and \(v\), we obtain
	\begin{equation}\label{eq:jacobian-difference-sde}
		\begin{cases}
			dD_t = A_t(u)D_t\,dt + \sum_{\ell=1}^{r}C_{\ell,t}(u)D_t\,dB_t^\ell + F_t\,dt +
			\sum_{\ell=1}^{r}G_{\ell,t}\,dB_t^\ell, \\[1mm]
			D_0=0,                                  \\[1mm]
		\end{cases}
	\end{equation}
	where
	\begin{equation*}
		F_t := \bigl(A_t(u)-A_t(v)\bigr)J_t(v)
	\end{equation*}
	and
	\begin{equation*}
		G_{\ell,t} := \bigl(C_{\ell,t}(u)-C_{\ell,t}(v)\bigr)J_t(v), \qquad \ell=1,\dots,r.
	\end{equation*}

	% using the assumptions
	For every \(M\in\mathbb R^{d\times d}\),
	\begin{equation*}
		\langle M,A_t(u)M\rangle_{\HS} \le L_x\|M\|_{\HS}^2,
	\end{equation*}
	and
	\begin{equation*}
		\sum_{\ell=1}^{r} \|C_{\ell,t}(u)M\|_{\HS}^2 \le L_\sigma^2\|M\|_{\HS}^2.
	\end{equation*}
	Indeed, both inequalities follow by applying \eqref{eq:basic-derivative-bounds} to the columns of \(M\).

	Similarly, applying \eqref{eq:spatial-lipschitz-derivatives} columnwise gives
	\begin{equation}\label{eq:forcing-coefficient-bounds}
		\|F_t\|_{\HS} \le H|\Delta X_t|\,\|J_t(v)\|_{\HS}
	\end{equation}
	and
	\begin{equation*}
		\left( \sum_{\ell=1}^{r}\|G_{\ell,t}\|_{\HS}^2 \right)^{1/2} \le H|\Delta X_t|\,\|J_t(v)\|_{\HS}.
	\end{equation*}
	Consequently, H\"older's inequality, \eqref{eq:flow-increment-moment-2m}, and \eqref{eq:jacobian-moment} imply
	\begin{align*}
		 & \E\|F_t\|_{\HS}^m + \E\left( \sum_{\ell=1}^{r}\|G_{\ell,t}\|_{\HS}^2 \right)^{m/2}                 \\
		 & \quad\le C_m \left( \E|\Delta X_t|^{2m} \right)^{1/2} \left( \E\|J_t(v)\|_{\HS}^{2m} \right)^{1/2} \\
		 & \quad\le C_me^{c_mt}|u-v|^m.
	\end{align*}
	Here we also used
	\(\|M\|_{\HS}\le\sqrt d\,\|M\|_{\op}\).

	Applying \Cref{lem:inhomogeneous-linear-moment} to \eqref{eq:jacobian-difference-sde}, and using \(D_0=0\), gives
	\begin{align*}
		\E\|D_t\|_{\HS}^m
		 & \le
		C_m|u-v|^m
		\int_0^t
		e^{c_m(t-s)}e^{c_ms}\,ds \\
		 & \le
		C_me^{c_mt}|u-v|^m,
	\end{align*}
	after enlarging \(c_m\). Therefore,
	\begin{equation}\label{eq:jacobian-increment}
		\E\|J_t(u)-J_t(v)\|_{\HS}^m \le C_me^{c_mt}|u-v|^m, \qquad u,v\in\mathbb R^d.
	\end{equation}
	In particular, the same estimate holds with the operator norm in place of the Hilbert--Schmidt norm.

	\paragraph{Step 3: Kolmogorov--Totoki and the compact supremum.}

	% define the cube???
	Let \(Q\subset\mathbb R^d\) be a compact cube containing \(K\), and choose \(u_0\in K\). By \eqref{eq:jacobian-increment}, \( u,v\in Q \)
	\begin{equation*}
		\E\|J_t(u)-J_t(v)\|_{\HS}^m \le C_me^{c_mt}|u-v|^m.
	\end{equation*}
	Since \(m>d\), this may be written as
	\begin{equation*}
		\E\|J_t(u)-J_t(v)\|_{\HS}^m \le C_me^{c_mt}|u-v|^{d+(m-d)}.
	\end{equation*}

	% application of Kolmogorov--Totoki theorem
	Hence the quantitative Kolmogorov--Totoki theorem, applied to the \(\mathbb R^{d\times d}\)-valued random field
	\begin{equation*}
		u\longmapsto J_t(u)-J_t(u_0),
	\end{equation*}
	gives
	\begin{equation}\label{eq:kolmogorov-totoki-bound}
		\E\left[\sup_{u\in Q} \|J_t(u)-J_t(u_0)\|_{\HS}^m \right] \le C_{m,Q}e^{c_mt}.
	\end{equation}
	Here we identify \(\mathbb R^{d\times d}\) with \(\mathbb R^{d^2}\).

	% strange paragraph
	Since \Cref{thm:jacobian-sde} already provides a spatially continuous version of \(J_t\), the continuous modification furnished by the
	Kolmogorov--Totoki theorem agrees with this version on \(Q\) almost surely.

	% final triangle inequality
	For every \(u\in K\),
	\begin{align*}
		\|J_t(u)\|_{\op} & \le \|J_t(u_0)\|_{\op} + \|J_t(u)-J_t(u_0)\|_{\HS}.
	\end{align*}
	% the finishing proof
	Consequently,
	\begin{equation*}
		\mathcal K_t(\cdot;K) \le \|J_t(u_0)\|_{\op} + \sup_{u\in Q} \|J_t(u)-J_t(u_0)\|_{\HS}.
	\end{equation*}
	Using \eqref{eq:jacobian-moment}, \eqref{eq:kolmogorov-totoki-bound}, and \((a+b)^m\le 2^{m-1}(a^m+b^m)\), we obtain
	\begin{equation*}
		\E\mathcal K_t(\cdot;K)^m \le C_{m,K}e^{c_mt}.
	\end{equation*}
	Finally,
	\begin{equation*}
		\E\mathcal K_t(\cdot;K)^p\le \left( \E\mathcal K_t(\cdot;K)^m \right)^{p/m} \le C_{p,K}e^{c_pt}.
	\end{equation*}
	This proves \eqref{eq:kt-exp-growth}.
\end{proof}

\clearpage

% ============================================================
% The averaged Jacobian
% ============================================================
\section{The averaged Jacobian}\label{sec:averaged-jacobian}

The preceding pathwise estimates use the quantity \(\mathcal K_t(\omega;K)\) which controls the worst possible stretching of the flow on the
compact set \(K\). Consider instead the averaged Jacobian for the transported measure \(\mu_t=\mu_0\circ T_t^{-1}\),
\begin{equation*}
	\mathcal E_t := \int_{\R^d}\|J_t(u)\|_{\HS}^2\,\mu_0(du).
\end{equation*}
It measures average deformation of the flow over the initial mass, rather than
the maximal deformation on the whole support.

\begin{proposition}[Averaged Jacobian energy estimate]\label{prop:averaged-jacobian-energy}
	Assume that the flow is differentiable in the initial label and that there
	exist constants \(L_{\mathrm{sym}},\Gamma<\infty\) such that
	\begin{equation}\label{eq:one-sided-drift-jacobian}
		\langle z,\partial_xa(x,\mu)z\rangle
		\le
		L_{\mathrm{sym}}|z|^2,
		\qquad x,z\in\R^d,\ \mu\in\cP_2(\R^d),
	\end{equation}
	and
	\begin{equation}\label{eq:quadratic-diffusion-jacobian}
		\sum_{k=1}^d|\partial_x\sigma^{(k)}(x)z|^2
		\le
		\Gamma^2|z|^2,
		\qquad x,z\in\R^d.
	\end{equation}
	Then
	\begin{equation}\label{eq:averaged-jacobian-bound}
		\E\mathcal E_t
		\le
		d\exp\{(2L_{\mathrm{sym}}+\Gamma^2)t\}.
	\end{equation}
\end{proposition}

\begin{proof}
	Recall that
	\begin{equation*}
		dJ_t(u) = A_t(u)J_t(u)\,dt + \sum_{k=1}^d C_t^{(k)}(u)J_t(u)\,dB_t^k,
	\end{equation*}
	where
	\begin{equation*}
		A_t(u)=\partial_xa(X_t(u),\mu_t), \qquad C_t^{(k)}(u)=\partial_x\sigma^{(k)}(X_t(u)).
	\end{equation*}
	Applying It\^o's formula to \(\|J_t(u)\|_{\HS}^2\) gives
	\begin{align*}
		d\|J_t(u)\|_{\HS}^2 & =
		\bigg[2\langle J_t(u),A_t(u)J_t(u)\rangle_{\HS} +
		\sum_{k=1}^d\|C_t^{(k)}(u)J_t(u)\|_{\HS}^2 \bigg]dt                                                 \\
		                    & \quad + 2\sum_{k=1}^d \langle J_t(u),C_t^{(k)}(u)J_t(u)\rangle_{\HS}\,dB_t^k.
	\end{align*}
	Since
	\begin{equation*}
		\langle J,A_tJ\rangle_{\HS} = \sum_{i=1}^d\langle Je_i,A_tJe_i\rangle,
	\end{equation*}
	condition \cref{eq:one-sided-drift-jacobian} gives
	\begin{equation*}
		2\langle J,A_tJ\rangle_{\HS} \le 2L_{\mathrm{sym}}\|J\|_{\HS}^2.
	\end{equation*}
	Similarly, \cref{eq:quadratic-diffusion-jacobian} gives
	\begin{equation*}
		\sum_{k=1}^d\|C_t^{(k)}J\|_{\HS}^2 = \sum_{i=1}^d\sum_{k=1}^d|C_t^{(k)}Je_i|^2 \le \Gamma^2\|J\|_{\HS}^2.
	\end{equation*}
	Integrating with respect to \(\mu_0(du)\) and taking expectations removes the
	martingale term, hence
	\begin{equation*}
		\frac{d}{dt}\E\mathcal E_t \le (2L_{\mathrm{sym}}+\Gamma^2)\E\mathcal E_t.
	\end{equation*}
	Since \(J_0(u)=I\), we have \(\mathcal E_0=\|I\|_{\HS}^2=d\). Gronwall's
	lemma proves \cref{eq:averaged-jacobian-bound}.
\end{proof}

To estimate \(\E\sup_{u\in K}\|J_t(u)\|^p\), one must control increments \(J_t(u)-J_t(v)\), which requires assumptions on the spatial regularity
of \(\partial_xa\) and \(\partial\sigma\). The averaged quantity \(\mathcal E_t\) does not compare two labels. It only estimates each \(J_t(u)\)
and integrates in \(u\), so the one-sided bounds \cref{eq:one-sided-drift-jacobian,eq:quadratic-diffusion-jacobian} are sufficient. \clearpage

% ============================================================
% Internal fluctuations
% ============================================================
\section{Internal fluctuations}\label{sec:internal-fluctuations}

% Motivation for this section
This section concerns a separate kernel-type interacting model. In contrast to \eqref{eq:main-sdewi}, both the drift and diffusion may depend on
the order parameter \(\Theta_t\). Consequently, the assumptions of \cref{sec:jacobian-bound,sec:averaged-jacobian} are not imposed here.

\begin{equation}\label{eq:kernel-sdewi}
	\begin{cases}
		dX_t(u) & = \bigg[ b(X_t(u),\Theta_t) + \int_{\R^d} K(X_t(u),X_t(v),\Theta_t)\,\mu_0(dv) \bigg]dt \\
		        & \quad+ \sum_{\ell=1}^m \sigma_\ell(X_t(u),\Theta_t)\,dB_t^\ell,                         \\[1mm]
		X_0(u)  & =u,\qquad u\in\R^d,                                                                     \\[1mm]
		\mu_t   & =\mu_0\circ X_t^{-1}.
	\end{cases}
\end{equation}
where
\begin{equation*}
	\theta(\mu):=\int_{\R^d}\Phi(x)\,\mu(dx),
\end{equation*}

% Dissipativity Assumptions on global Lipschitz and growth hypotheses.
\begin{assumption}[Well-posedness of \Cref{eq:kernel-sdewi}]\label{ass:kernel-wellposedness}
	The maps \(b,K,\sigma_\ell,\Phi\) are globally Lipschitz and have at most linear growth.
	More explicitly, there exists \(L<\infty\) such that
	\begin{align*}
		|\Phi(x)-\Phi(y)|           & \le L|x-y|,                               \\
		|b(x,\theta)-b(y,\eta)|     & \le L\big(|x-y|+|\theta-\eta|\big),       \\
		|K(x,z,\theta)-K(y,w,\eta)| & \le L\big(|x-y|+|z-w|+|\theta-\eta|\big),
	\end{align*}
	and
	\begin{equation*}
		\sum_{\ell=1}^m |\sigma_\ell(x,\theta)-\sigma_\ell(y,\eta)|^2 \le L^2\big(|x-y|^2+|\theta-\eta|^2\big).
	\end{equation*}
\end{assumption}

% Well-posedness of the SDEWI
Under \Cref{ass:kernel-wellposedness}, equation \eqref{eq:kernel-sdewi} admits a unique global continuous solution.

This section studies the pairwise energy
\begin{equation*}
	\mathcal D_t := \iint_{\R^d\times\R^d} |X_t(u)-X_t(v)|^2\,\mu_0(du)\mu_0(dv).
\end{equation*}

% ============================================================
% Dissipativity assumptions
% ============================================================
\subsection{Dissipativity assumptions}\label{subsec:dissipativity-assumptions}
Equivalently, if
\begin{equation*}
	\bar X_t:=\int_{\R^d}X_t(u)\,\mu_0(du),
\end{equation*}
and
\begin{equation*}
	\mathcal V_t := \int_{\R^d}|X_t(u)-\bar X_t|^2\,\mu_0(du),
\end{equation*}
then
\begin{equation*}
	\mathcal D_t=2\mathcal V_t.
\end{equation*}
Indeed,
\begin{align*}
	\mathcal D_t & = \iint |X_t(u)-X_t(v)|^2\,\mu_0(du)\mu_0(dv)                                                \\
	             & = \iint \big( |X_t(u)|^2+|X_t(v)|^2 -2\langle X_t(u),X_t(v)\rangle \big)\,\mu_0(du)\mu_0(dv) \\
	             & = 2\int |X_t(u)|^2\,\mu_0(du)- 2\left|\int X_t(u)\,\mu_0(du)\right|^2                        \\
	             & = 2\int |X_t(u)-\bar X_t|^2\,\mu_0(du)= 2\mathcal V_t.
\end{align*}
We assume the following dissipativity condition.
\begin{assumption}[Pairwise dissipativity]\label{ass:pairwise-dissipativity}
	There exists \(\lambda>0\) such that for every \(x,y\in\R^d\), every \(\theta\in\R^q\), and every
	\(\nu\in\cP_2(\R^d)\),
	\begin{align*}
		 & 2\left\langle x-y, b(x,\theta)-b(y,\theta) + \int_{\R^d} \big[K(x,z,\theta)-K(y,z,\theta)\big] \nu(dz) \right\rangle \\
		 & \qquad + \sum_{\ell=1}^m |\sigma_\ell(x,\theta)-\sigma_\ell(y,\theta)|^2 \le -\lambda |x-y|^2.
	\end{align*}
\end{assumption}

The dissipativity condition covers several basic models.

\begin{example}[Ornstein--Uhlenbeck SDEWI]
	Consider
	\begin{equation*}
		dX_t(u)= -\lambda_0\big(X_t(u)-m_t\big)\,dt + \sigma(X_t(u))\,dB_t, \qquad m_t=\int_{\R^d} X_t(v)\,\mu_0(dv).
	\end{equation*}
	If
	\begin{equation*}
		\|\sigma(x)-\sigma(y)\|_{\HS} \le L_\sigma |x-y|,
	\end{equation*}
	then
	\begin{equation*}
		\E\mathcal D_t \le e^{-(2\lambda_0-L_\sigma^2)t}\mathcal D_0.
	\end{equation*}
	Thus the cloud contracts whenever \(2\lambda_0>L_\sigma^2\).
\end{example}

\begin{example}[Gradient systems]
	Take
	\begin{equation*}
		b(x,\theta)=-\nabla V(x), \qquad K(x,y,\theta)=-\nabla W(x-y).
	\end{equation*}
	Then
	\begin{equation*}
		dX_t(u) =  \left[ -\nabla V(X_t(u)) - \int_{\R^d} \nabla W(X_t(u)-X_t(v))\,\mu_0(dv) \right]dt + \sigma(X_t(u))\,dB_t.
	\end{equation*}
	If
	\begin{equation*}
		D^2V\ge \rho I, \qquad D^2W\ge \kappa I, \qquad \|\sigma(x)-\sigma(y)\| \le L_\sigma |x-y|,
	\end{equation*}
	then
	\begin{equation*}
		\E\mathcal D_t \le e^{-(2\rho+2\kappa-L_\sigma^2)t}\mathcal D_0.
	\end{equation*}
\end{example}

\begin{example}[Nonlinear statistics]
	Let
	\begin{equation*}
		\Theta_t=\int_{\R^d} \psi(X_t(v))\,\mu_0(dv),
	\end{equation*}
	and consider
	\begin{equation*}
		dX_t(u) = -\lambda_0\big(X_t(u)-\Theta_t\big)\,dt + \sigma(X_t(u),\Theta_t)\,dB_t.
	\end{equation*}
	For two labels \(u,v\), the order parameter cancels in the drift
	difference:
	\begin{equation*}
		-\lambda_0\big(X_t(u)-\Theta_t\big) + \lambda_0\big(X_t(v)-\Theta_t\big) = -\lambda_0\big(X_t(u)-X_t(v)\big).
	\end{equation*}

\end{example}

% ============================================================
% Proof
% ============================================================
\subsection{Proof}\label{subsec:synchronization-proof}
\begin{theorem}[Synchronization under dissipativity assumptions]\label{thm:synchronization-dissipativity}
	Assume that \(\mu_0\in\cP_2(\R^d)\) and that \Cref{ass:pairwise-dissipativity} holds. Then
	\begin{equation*}
		\E\mathcal D_t \le e^{-\lambda t}\mathcal D_0.
	\end{equation*}

	In particular,
	\begin{equation*}
		\E W_2^2(\mu_t,\delta_{\bar X_t}) \le e^{-\lambda t}W_2^2(\mu_0,\delta_{\bar X_0}).
	\end{equation*}
	Thus the transported measure \(\mu_t\) concentrates around its random center of mass \(\bar X_t\).
\end{theorem}

% Proof idea:
% 1. Differentiate the stochastic flow.
\begin{proof}
	% Proof roadmap: first prove pairwise contraction, then integrate over labels.
	Fix two labels \(u,v\in\R^d\). Define the difference
	\begin{equation*}
		Z_t^{u,v}:=X_t(u)-X_t(v).
	\end{equation*}
	We first prove that under \Cref{ass:pairwise-dissipativity}
	\begin{equation*}
		\E|Z_t^{u,v}|^2 \le e^{-\lambda t}|u-v|^2,
	\end{equation*}
	then we integrate this estimate with respect to \(\mu_0(du)\mu_0(dv)\).

	Writing \cref{eq:kernel-sdewi} for \(X_t(u)\) and for \(X_t(v)\) and then subtracting the two equations gives
	\begin{equation*}
		dZ_t^{u,v} = G_t^{u,v}\,dt + \sum_{\ell=1}^m H_{\ell,t}^{u,v}\,dB_t^\ell,
	\end{equation*}
	where the drift coefficient
	\begin{align*}
		G_t^{u,v} & := b(X_t(u),\Theta_t)-b(X_t(v),\Theta_t)                                                   \\
		          & +  \int_{\R^d} \big[ K(X_t(u),X_t(w),\Theta_t) -K(X_t(v),X_t(w),\Theta_t)\big]\,\mu_0(dw),
	\end{align*}
	and the diffusion coefficient is
	\begin{equation*}
		H_{\ell,t}^{u,v} := \sigma_\ell(X_t(u),\Theta_t) - \sigma_\ell(X_t(v),\Theta_t).
	\end{equation*}

	Since \(\mu_t =\mu_0\circ X_t^{-1}\), we may rewrite the interaction term as
	\begin{equation*}
		\int_{\R^d}K(X_t(u),X_t(w),\Theta_t)\,\mu_0(dw) = \int_{\R^d}K(X_t(u),z,\Theta_t)\,\mu_t(dz).
	\end{equation*}
	Thus
	\begin{align*}
		G_t^{u,v} & := b(X_t(u),\Theta_t)-b(X_t(v),\Theta_t)                                          \\
		          & + \int_{\R^d} \big[ K(X_t(u),z,\Theta_t) - K(X_t(v),z,\Theta_t) \big]\,\mu_t(dz).
	\end{align*}

	Applying It\^o's formula to \(|Z_t^{u,v}|^2\), we obtain
	\begin{equation*}
		d|Z_t^{u,v}|^2 = 2\langle Z_t^{u,v},G_t^{u,v}\rangle\,dt + \sum_{\ell=1}^m |H_{\ell,t}^{u,v}|^2\,dt + 2\sum_{\ell=1}^m \langle Z_t^{u,v},H_{\ell,t}^{u,v}\rangle\,dB_t^\ell.
	\end{equation*}
	Apply \Cref{ass:pairwise-dissipativity} with
	\(x=X_t(u)\), \(y=X_t(v)\), \(\theta=\Theta_t\), and \(\nu=\mu_t\). Then
	\begin{equation*} 2\langle Z_t^{u,v},G_t^{u,v}\rangle +  \sum_{\ell=1}^m |H_{\ell,t}^{u,v}|^2 \le -\lambda |Z_t^{u,v}|^2.
	\end{equation*}
	Therefore
	\begin{equation}\label{eq:pairwise-contraction-differential}
		d|Z_t^{u,v}|^2 \le -\lambda |Z_t^{u,v}|^2\,dt + 2\sum_{\ell=1}^m \langle Z_t^{u,v},H_{\ell,t}^{u,v}\rangle\,dB_t^\ell.
	\end{equation}

	Set the new variable
	\begin{equation*}
		Y_t^{u,v}:=e^{\lambda t}|Z_t^{u,v}|^2.
	\end{equation*}
	By It\^o's formula,
	\begin{align*}
		dY_t^{u,v} & = e^{\lambda t} \left[ \lambda |Z_t^{u,v}|^2 + 2\langle Z_t^{u,v},G_t^{u,v}\rangle + \sum_{\ell=1}^m |H_{\ell,t}^{u,v}|^2 \right]dt \\
		           & \quad+ 2e^{\lambda t} \sum_{\ell=1}^m \langle Z_t^{u,v},H_{\ell,t}^{u,v}\rangle\,dB_t^\ell.
	\end{align*}
	By the drift bound \cref{eq:pairwise-contraction-differential}, the drift term above is nonpositive. Hence
	\begin{equation*}
		dY_t^{u,v} \le 2e^{\lambda t} \sum_{\ell=1}^m \langle Z_t^{u,v},H_{\ell,t}^{u,v}\rangle\,dB_t^\ell.
	\end{equation*}
	% Thus \(Y_t^{u,v}\) is a nonnegative local supermartingale.

	To make this completely rigorous, introduce the localization sequence
	\begin{equation*}
		\tau_R := \inf\{t\ge0: |X_t(u)|+|X_t(v)|+|\Theta_t|\ge R\}.
	\end{equation*}
	Since the solution is global and continuous, \(\tau_R\uparrow\infty\) almost surely as \(R\to\infty\). Stopping this inequality at \(\tau_R\), we get
	\begin{equation*}
		Y_{t\wedge\tau_R}^{u,v} \le Y_0^{u,v} + 2\sum_{\ell=1}^m \int_0^{t\wedge\tau_R} e^{\lambda s} \langle Z_s^{u,v},H_{\ell,s}^{u,v}\rangle\,dB_s^\ell.
	\end{equation*}
	The stopped stochastic integral is a true martingale. Therefore, taking
	expectations gives
	\begin{equation*}
		\E Y_{t\wedge\tau_R}^{u,v} \le Y_0^{u,v}.
	\end{equation*}
	Since \(Y_0^{u,v}=|u-v|^2\), we have
	\begin{equation*}
		\E\left[ e^{\lambda(t\wedge\tau_R)} |Z_{t\wedge\tau_R}^{u,v}|^2 \right] \le |u-v|^2.
	\end{equation*}
	Letting \(R\to\infty\) and using Fatou's lemma yields
	\begin{equation*}
		\E\left[e^{\lambda t}|Z_t^{u,v}|^2\right] \le |u-v|^2.
	\end{equation*}
	Therefore
	\begin{equation*}
		\E|X_t(u)-X_t(v)|^2 \le  e^{-\lambda t}|u-v|^2.
	\end{equation*}
	This proves the pairwise synchronization estimate.

	Now integrate the pairwise synchronization estimate with respect to \(\mu_0(du)\mu_0(dv)\). Since \(\mu_0\in\cP_2(\R^d)\),
	\begin{equation*}
		\mathcal D_0 = \iint |u-v|^2\,\mu_0(du)\mu_0(dv)<\infty.
	\end{equation*}
	By Fubini's theorem,
	\begin{align*}
		\E\mathcal D_t & = \E \iint |X_t(u)-X_t(v)|^2\,\mu_0(du)\mu_0(dv)      \\
		               & = \iint \E|X_t(u)-X_t(v)|^2\,\mu_0(du)\mu_0(dv)       \\
		               & \le e^{-\lambda t} \iint |u-v|^2\,\mu_0(du)\mu_0(dv).
	\end{align*}
	Hence \(\E\mathcal D_t\le e^{-\lambda t}\mathcal D_0\), proving
	the first estimate in the theorem.

	Therefore
	\begin{equation*}
		\E\mathcal V_t = \frac12\E\mathcal D_t \le \frac12 e^{-\lambda t}\mathcal D_0 = e^{-\lambda t}\mathcal V_0.
	\end{equation*}
	This proves the internal fluctuation decay estimate.

	Finally, since \(\mu_t =\mu_0\circ X_t^{-1}\),
	\begin{equation*}
		\int_{\R^d}|x-\bar X_t|^2\,\mu_t(dx) = \int_{\R^d}|X_t(u)-\bar X_t|^2\,\mu_0(du) = \mathcal V_t.
	\end{equation*}
	For the Dirac measure \(\delta_{\bar X_t}\),
	\begin{equation*}
		W_2^2(\mu_t,\delta_{\bar X_t}) = \int_{\R^d}|x-\bar X_t|^2\,\mu_t(dx).
	\end{equation*}
	Therefore
	\begin{equation*}
		W_2^2(\mu_t,\delta_{\bar X_t})=\mathcal V_t.
	\end{equation*}
	Taking expectations and using the previous estimate gives
	\begin{equation*}
		\E W_2^2(\mu_t,\delta_{\bar X_t}) \le e^{-\lambda t}W_2^2(\mu_0,\delta_{\bar X_0}).
	\end{equation*}
\end{proof}

% The final result of the section
This result is the contractive counterpart of the distortion estimates in the preceding sections.
\Cref{sec:jacobian-bound,sec:averaged-jacobian} control the growth of infinitesimal perturbations, either uniformly or on average.
Here the pairwise dissipativity assumption controls finite separations
directly and forces their average to decay. Since \(\mathcal D_t=2\mathcal V_t\), this finite-scale contraction is
exactly the collapse of the transported measure around its random center of mass.

\bibliographystyle{apalike}
\bibliography{reference}

\end{document}
